Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-04-21 ms/LMA_method.Rmd
add subsection for Bayes

@@

The amount of intraspecific variation of LMA was calculated using coefficient of variation of log-normally distributed data; $CV = \sqrt{exp(s_{\mathrm{ln}}^2) - 1}$, where $s_{\mathrm{ln}}$ is the sample standard deviation in the log-transformed data [@Julious2000].
We also considered CV on the arithmetic scale; $CV = s/\bar{x}$, where $s$ is the sample standard deviation and $\bar{x}$ is the sample mean.
However, these two different CV calculations only negligibly altered the bivariate relationship and did not change our conclusion, and thus we will report CV of log-normally distributed data in the main text.

SMA analyses were run using the R package *smatr* [@Warton2012a]; this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 1000 bootstraps.
All statistical analyses were conducted in R version 4.1.2 [@RCoreTeam2021].

>>>>

*new subsection*

To quantify the effect of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA, we built hierarchical Bayes models.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf size and leaf thickness on the log-scale;

$$
\mathrm{ln}(LMA_{Wi}) \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i)
$$

$$
\tilde{\mu_i} = \beta_0 + \beta_1 \mathrm{ln}(LA_{i}) + \beta_2 \mathrm{ln}(LT_{i})
$$

where $LMA_{Wi}$, $LMA_{Di}$, $LA_{i}$, $LT_{i}$ are the whole-leaf LMA, leaf disc LMA, leaf area and leaf thickness for tree *i*, respectively.
$\tilde{\mu_i}$ is the effect of leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree *i*, and $\sigma_i$ is the variance of whole-leaf LMA for tree *i*, and $\beta$ are coefficients.

The variance of whole-leaf LMA ($\sigma_i$) was assumed to follow a normal distribution;

$$
\mathrm{ln}(\sigma_i) \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega)
$$

$$
\tilde{\sigma_i} = \gamma_0 + \gamma_1 \mathrm{ln}(LA_{i}) + \gamma_2 \mathrm{ln}(LT_{i})
$$

where $\tilde{\sigma_i}$ is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
$\omega$ is a scaling hyperparameter, and $\gamma$ are coefficients.
To allow comparisons among parameter estimates, $\mathrm{ln}(LA)$ and $\mathrm{ln}(LT)$ were scaled to a mean of 0 and standard deviation of 1.

Posterior distributions of all parameters were estimated using the Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan [@Carpenter2017] using the weakly-informative priors [@Gelman2008].
See supplement XXXX for more detail.
The Stan code use to fit models is available from Github at: [https://github.com/mattocci27/leaf-disc](https://github.com/mattocci27/leaf-disc).
Convergence of the posterior distribution was assessed with the Gelman-Rubin statistic with a convergence threshold of 1.1 for all parameters [@Gelman2013].

*new subsection*

@@

*new subsection?*

The amount of intraspecific variation (ITV) was calculated using Bao's CV estimator [@Yang2020];

$$
CV = CV_1 - \frac{CV_1^3}{N} + \frac{CV_1}{4N} + \frac{CV_1^2 \gamma_1}{2N} + \frac{CV_1 \gamma_2}{8N}
$$

where $CV_1 = s/\bar{x}$, $s$ is the sample standard deviation and $\bar{x}$ is the sample mean, *N* is the sample size, and $\gamma_1$ and $\gamma_2$ are the skewness and kurtosis of the trait distribution.
Bao's CV estimator is a robust method to quantify ITV based on coefficient of variation (CV) when sample size is small [@Yang2020].

SMA analyses were run using the R package *smatr* [@Warton2012a]; this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.
All statistical analyses were conducted in R version 4.1.3 [@RCoreTeam2022].

### 2022-04-21 ms/LMA_method.tex
add subsection for Bayes

\textbf{RESULTS:} Disc-based estimates of species and individual mean
LMA matched well whole-leaf estimates. Thin and large leaves had weaker
correlations between leaf disc and whole-leaf estimates. Variability of
LMA estimates within species (i.e., amount of intraspecific variation)
only weakly matched between whole-leaf and disc-based estimates
(\emph{R\textsuperscript{2}} = 0.32). We also found that sample with
small total dry mass of leaf discs tended to show greater differences
between leaf disc and whole-leaf LMA.

@@

@@

The amount of intraspecific variation of LMA was calculated using
coefficient of variation of log-normally distributed data;
\(CV = \sqrt{exp(s_{\mathrm{ln}}^2) - 1}\), where \(s_{\mathrm{ln}}\) is
the sample standard deviation in the log-transformed data
(\protect\hyperlink{ref-Julious2000}{Julious and Debarnot, 2000}). We
also considered CV on the arithmetic scale; \(CV = s/\bar{x}\), where
\(s\) is the sample standard deviation and \(\bar{x}\) is the sample
mean. However, these two different CV calculations only negligibly
altered the bivariate relationship and did not change our conclusion,
and thus we will report CV of log-normally distributed data in the main
text.

SMA analyses were run using the R package \emph{smatr}
(\protect\hyperlink{ref-Warton2012a}{Warton et al., 2012}); this
included tests for slope heterogeneity and elevation differences between
disc-based and whole-leaf estimates of LMA values. Confidence intervals
of SMA estimates were based on 1000 bootstraps. All statistical analyses
were conducted in R version 4.1.2
(\protect\hyperlink{ref-RCoreTeam2021}{R Core Team, 2021}).

@@

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values were smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships of the whole-leaf LMA : disc leaf LMA ratio
with total dry mass for the leaf disc (we weighted 2-3 discs together),
with leaf thickness, and with leaf area were heteroscedastic (unequal
variance, \emph{P} \textless{} 0.001; Appendix S3). Samples with small
total dry mass, thin leaves, and larger leaves tended to show greater
variance (Fig. 2; Appendix S3). The total dry mass of leaf disc was
positively correlated with both leaf thickness and leaf punch size
(Appendix S4). Amounts of intraspecific variation (CV; coefficient of
variation) only weakly matched between whole-leaf and disc-based
estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.32).

@@

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.32). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored in community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to show high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Julious2000}{}}%
Julious, S. A., and C. A. Debarnot. 2000. Why are pharmacokinetic data
summarized by arithmetic means? \emph{Journal of Biopharmaceutical
Statistics} 10: 55--71.

@@

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2021}{}}%
R Core Team. 2021. R: {A} language and environment for statistical
computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.

@@

@@

\begin{longtable}[]{@{}lllr@{}}
\toprule
Data & Slope & Intercept & \emph{R\textsuperscript{2}} \\
\midrule
\endhead
All & 0.97 {[}0.94, 1.01{]} & \textbf{0.09 {[}0.02, 0.16{]}} & 0.87 \\
Thick-large & 0.93 {[}0.85, 1.03{]} & \textbf{0.19 {[}0.01, 0.37{]}} &
0.86 \\
Thick-small & 1.02 {[}0.97, 1.07{]} & 0 {[}-0.11, 0.11{]} & 0.93 \\
Thin-large & 0.91 {[}0.82, 1.02{]} & \textbf{0.2 {[}0.02, 0.37{]}} &
0.71 \\
Thin-small & 0.93 {[}0.83, 1.03{]} & 0.16 {[}-0.02, 0.35{]} & 0.84 \\
\bottomrule
\end{longtable}

>>>>

\textbf{RESULTS:} Disc-based estimates of species and individual mean
LMA matched well whole-leaf estimates. Thin and large leaves had weaker
correlations between leaf disc and whole-leaf estimates. Variability of
LMA estimates within species (i.e., amount of intraspecific variation)
only weakly matched between whole-leaf and disc-based estimates
(\emph{R\textsuperscript{2}} = 0.33). We also found that sample with
small total dry mass of leaf discs tended to show greater differences
between leaf disc and whole-leaf LMA.

@@

\emph{new subsection}

To quantify the effect of leaf sizes and thicknesses on the relationship
between whole-leaf and leaf disc LMA, we built hierarchical Bayes
models. The whole-leaf LMA was assumed to follow a normal distribution
(N) with the mean of leaf disc LMA and covariates of leaf size and leaf
thickness on the log-scale;

\[
\mathrm{ln}(LMA_{Wi}) \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i)
\]

\[
\tilde{\mu_i} = \beta_0 + \beta_1 \mathrm{ln}(LA_{i}) + \beta_2 \mathrm{ln}(LT_{i})
\]

where \(LMA_{Wi}\), \(LMA_{Di}\), \(LA_{i}\), \(LT_{i}\) are the
whole-leaf LMA, leaf disc LMA, leaf area and leaf thickness for tree
\emph{i}, respectively. \(\tilde{\mu_i}\) is the effect of leaf area and
leaf thickness on the mean estimates for the whole-leaf LMA of tree
\emph{i}, and \(\sigma_i\) is the variance of whole-leaf LMA for tree
\emph{i}, and \(\beta\) are coefficients.

The variance of whole-leaf LMA (\(\sigma_i\)) was assumed to follow a
normal distribution;

\[
\mathrm{ln}(\sigma_i) \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega)
\]

\[
\tilde{\sigma_i} = \gamma_0 + \gamma_1 \mathrm{ln}(LA_{i}) + \gamma_2 \mathrm{ln}(LT_{i})
\]

where \(\tilde{\sigma_i}\) is the effect of leaf area and leaf thickness
on the variance of whole-leaf LMA, \(\omega\) is a scaling
hyperparameter, and \(\gamma\) are coefficients. To allow comparisons
among parameter estimates, \(\mathrm{ln}(LA)\) and \(\mathrm{ln}(LT)\)
were scaled to a mean of 0 and standard deviation of 1.

Posterior distributions of all parameters were estimated using the
Hamiltonian Monte Carlo algorithm (HMC) implemented in Stan
(\protect\hyperlink{ref-Carpenter2017}{Carpenter et al., 2017}) using
the weakly-informative priors (\protect\hyperlink{ref-Gelman2008}{Gelman
et al., 2008}). See supplement XXXX for more detail. The Stan code use
to fit models is available from Github at:
\url{https://github.com/mattocci27/leaf-disc}. Convergence of the
posterior distribution was assessed with the Gelman-Rubin statistic with
a convergence threshold of 1.1 for all parameters
(\protect\hyperlink{ref-Gelman2013}{Gelman et al., 2013}).

\emph{new subsection}

@@

\emph{new subsection?}

The amount of intraspecific variation (ITV) was calculated using Bao's
CV estimator (\protect\hyperlink{ref-Yang2020}{Yang et al., 2020});

\[
CV = CV_1 - \frac{CV_1^3}{N} + \frac{CV_1}{4N} + \frac{CV_1^2 \gamma_1}{2N} + \frac{CV_1 \gamma_2}{8N}
\]

where \(CV_1 = s/\bar{x}\), \(s\) is the sample standard deviation and
\(\bar{x}\) is the sample mean, \emph{N} is the sample size, and
\(\gamma_1\) and \(\gamma_2\) are the skewness and kurtosis of the trait
distribution. Bao's CV estimator is a robust method to quantify ITV
based on coefficient of variation (CV) when sample size is small
(\protect\hyperlink{ref-Yang2020}{Yang et al., 2020}).

SMA analyses were run using the R package \emph{smatr}
(\protect\hyperlink{ref-Warton2012a}{Warton et al., 2012}); this
included tests for slope heterogeneity and elevation differences between
disc-based and whole-leaf estimates of LMA values. Confidence intervals
of SMA estimates were based on 2000 bootstraps. All statistical analyses
were conducted in R version 4.1.3
(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}).

@@

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values were smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships of the whole-leaf LMA : disc leaf LMA ratio
with total dry mass for the leaf disc (we weighted 2-3 discs together),
with leaf thickness, and with leaf area were heteroscedastic (unequal
variance, \emph{P} \textless{} 0.001; Appendix S3). Samples with small
total dry mass, thin leaves, and larger leaves tended to show greater
variance (Fig. 2; Appendix S3). The total dry mass of leaf disc was
positively correlated with both leaf thickness and leaf punch size
(Appendix S4). Amounts of intraspecific variation (CV; coefficient of
variation) only weakly matched between whole-leaf and disc-based
estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.33).

@@

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.33). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored in community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to show high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Carpenter2017}{}}%
Carpenter, B., A. Gelman, M. D. Hoffman, D. Lee, B. Goodrich, M.
Betancourt, M. Brubaker, et al. 2017.
\href{https://doi.org/10.18637/jss.v076.i01}{Stan : {A Probabilistic
Programming Language}}. \emph{Journal of Statistical Software} 76:
1--32.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2013}{}}%
Gelman, A., J. B. Carlin, H. S. Stern, D. B. Dunson, A. Vehtari, and D.
B. Rubin. 2013. Bayesian {Data Analysis}, {Third Edition}. {Taylor \&
Francis}.

\leavevmode\vadjust pre{\hypertarget{ref-Gelman2008}{}}%
Gelman, A., A. Jakulin, M. G. Pittau, and Y. S. Su. 2008.
\href{https://doi.org/10.1214/08-AOAS191}{A weakly informative default
prior distribution for logistic and other regression models}.
\emph{Annals of Applied Statistics} 2: 1360--1383.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
R Core Team. 2022. R: {A} language and environment for statistical
computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yang2020}{}}%
Yang, J., J. Lu, Y. Chen, E. Yan, J. Hu, X. Wang, and G. Shen. 2020.
Large {Underestimation} of {Intraspecific Trait Variation} and {Its
Improvements}. \emph{Frontiers in Plant Science} 11.

@@

\begin{longtable}[]{@{}lllr@{}}
\toprule()
Data & Slope & Intercept & \emph{R\textsuperscript{2}} \\
\midrule()
\endhead
All & 0.97 {[}0.94, 1.01{]} & \textbf{0.09 {[}0.02, 0.16{]}} & 0.87 \\
Thick-large & 0.93 {[}0.85, 1.03{]} & \textbf{0.19 {[}0.01, 0.37{]}} &
0.86 \\
Thick-small & 1.02 {[}0.97, 1.07{]} & 0 {[}-0.11, 0.11{]} & 0.93 \\
Thin-large & 0.91 {[}0.82, 1.02{]} & \textbf{0.2 {[}0.02, 0.37{]}} &
0.71 \\
Thin-small & 0.93 {[}0.83, 1.03{]} & 0.16 {[}-0.02, 0.35{]} & 0.84 \\
\bottomrule()
\end{longtable}
